\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage{hyperref}
\onehalfspacing



\begin{document}

\title{Pol Sci 630: Problem Set 7 - Short Research Proposal}

\author{Prepared by: Jan Vogler (\href{mailto:jan.vogler@duke.edu}{jan.vogler@duke.edu})}

\date{Due Date: Thursday, October 15th, 2015, 10 AM (Beginning of Class)}
 
\maketitle 



Please use a *single* PDF file created through knitr to submit your answers.

Make sure that the PDF files you submit do not include any references to your identity. The grading will happen anonymously. You can submit your answer at the following website: \url{http://ps630-f15.herokuapp.com/}



\section*{Short Research Proposal}

\paragraph{Please write a short research proposal with a (maximum) length of three paragraphs.}

\subparagraph{a)} Write one paragraph on the background of your research idea. What is the question you want to ask? How does it matter? Which academic works are most closely related to the specific question you want to answer?

\subparagraph{b)} Write one paragraph on how the theoretical concepts you are interested in can be measured as variables. What is your unit of analysis? What is your dependent variable? What is your independent variable? What are your control variables?

\subparagraph{c)} Write one paragraph on which datasets would be appropriate to use. Which data sets contain the data you need? How many observations are available over which time frame? How difficult would it be to obtain the data?



\end{document}