---
title: "Tutorial 11: Simulations and autocorrelation across time and space"
author: "Jan Vogler (jan.vogler@duke.edu)"
date: "November 6, 2015"
output: pdf_document
---

# Today's Agenda

1. Model simulations
2. Using model simulations to make predictions
3. Estimating interaction effects through simulations
4. Autocorrelation across time and space

**Credit to Professor Chris Johnston. The code for the simulation approach introduced here is from his class.**



# 1. Model simulations

We will look at data from a data set that we have not used previously, data on attitudes toward the United States Supreme Court.

```{r}
setwd("C:/Users/Jan/OneDrive/Documents/GitHub/ps630_lab/w11/")
courtdata <- read.table("courtdata.txt", header=TRUE)
summary(courtdata)
```

What is the meaning of the variables in the data set?

## Independent Variables

college = dummy for college education

ideo = 5-point ideological self-identification, ranging from "very liberal" to "very conservative"

soph = 10-item political knowledge scale

black, hispanic = dummies for respective categories

income = 15-point household income category scale

ruling = dummy indicating respondent correctly identified the Affordable Care Act ruling

male = dummy indicating male gender

## Dependent Variable

We are interested in attitudes of US citizens toward the Supreme Court. For this purpose we look at the variable "sclaw" ("Supreme Court Law"). The variable was based on the following statement:

"The Supreme Court should be allowed to throw out any law it deems unconstitutional"

The variable ranges from "strongly agree" to "strongly disagree".

Let us estimate a linear model.

```{r}
model1 <- lm(sclaw ~ ruling + ideo + soph + age + male + black + hisp + college + income, data=courtdata)
summary(model1)
```

When we estimate a linear model, all our coefficients are assumed to be normally distributed random variables with a mean and a variance that depends on the data. We can make use of this fact by simulating different versions of the model.

For the simulations we need the package "arm". Please make sure to install it via the following command: install.packages("arm")

```{r}
library(arm)
model1.sims <- sim(model1, n.sims=1000)
```



# 2. Using model simulations to make predictions

```{r}

```



# 3. Estimating interaction effects through simulations

```{r}
```



# 4. Autocorrelation across time and space

```{r}
```
